On Sample Size Needed for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates

### 2024-01-08 Manuscript/manuscript.tex
fixed differences between distributions and block sizes

\subsection*{Normal Marginal Distribution}
%We designed a simulation study to compare the performance of the different block 
%bootstrap CI methods. In particular, 
We generated time series $X_t$ from a 1st 
order autoregressive (AR(1)) process:
\begin{equation*}
X_t = \phi X_{t-1} + \epsilon_t,
\end{equation*}
where $\phi$ is an autoregressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: $\{-0.4, -0.2, 0.0, 0.2, 0.4\}$. We only
used serial dependences as strong as 0.4, because we only seek to 
establish the 
general trend as the strength of the autocorrelation 
increases, and how it varies depending on the sign of the autocorrelation and 
the parameter of interest. The 
series $X_t$ has mean zero and variance 
$\sigma_x^2 = (1 - \phi^2) \sigma_{\epsilon}^2$, so for each value of $\phi$, we 
set $\sigma_{\epsilon}^2 = 1 / (1 - \phi^2)$ such that $\sigma_x^2 = 1$.

@@

The coverage rates of the CIs were used to compare the performance of CIs. Let
$\hat\theta_{L, r}$ and $\hat\theta_{U, r}$ be the lower and upper bound,
respectively, for the confidence interval constructed for each replicate
$r \in \{1, \ldots, R\}$, where $R$ is the number of replicates. Then
$cov = \#\{\hat\theta_{L, r} < \theta < \hat\theta_{U, r} \}/R$, for
$r \in \{1, \ldots, R\}$. If a CI method is valid, then the coverage rate is
expected to match the nominal level. Because it is unlikely for the coverage to 
exactly match the nominal level, we can construct a 95\% Wald-type CI of the 
coverage rate 
(which is an estimate of a proportion with $R = 10000$). If the proportion 0.95 
is included in the interval, the block bootstrap method is likely performing 
well. If all values in the interval are below 0.95, the results would suggest 
that the method either is providing inaccurate estimation, is underestimating 
the process' variability, or a combination of both. If all values in the 
interval are above .95, the results suggest that the method is overestimating 
the process' variability. \citet{brown2001interval} notes that the Wald-type 
interval has poor coverage as the proportion approaches 0 or 1.
\textbf{Figures~\ref{fig:mu1}--\ref{fig:phi1}} 
summarizes the empirical coverage rates and the 95\% confidence intervals of the 
real coverage for a standard normal marginal distribution using block bootstrap
with $l = \lceil n^{1/3} \rceil$, generated using the 
R \textsl{ggplot2} package.\citep{ggplot2}

\subsection*{Exponential Marginal Distribution}
Additionally, we evaluated the performance of block bootstrap when used to 
estimate
the mean, standard deviation, and the lag-1 autocorrelation coefficient of a
stationary series with an exponential marginal distribution. The motivation of
for this study was to investigate if the results are robust to nonnormal
marginal distributions.

@@

\textbf{Figures~\ref{fig:exp_mu1}--\ref{fig:exp_phi1}} 
summarizes the empirical coverage rates and the 95\% confidence intervals of the 
real coverage for a exponential marginal distribution using block bootstrap
with $l = \lceil n^{1/3} \rceil$.

@@

@@

\subsection*{Normal Marginal Distribution}
For estimating the mean parameter~$\mu$, \textbf{Figure~\ref{fig:mu1}} suggests 
that all methods eventually approach correct coverage of $\mu$ as sample size 
increases. Student's $t$ CIs appear to need the smallest sample size to achieve 
correct coverage, except for samples with strong negative dependence, in which 
case, they actually over-cover $\mu$ for smaller sample sizes. For instance, for 
a sample with $n = 100$ and $\phi = -0.4$, the lower bound for a Students $t$ 
CI's coverage of $\mu$ is greater than 95\%, whereas the coverage intervals for 
other methods contain 95\%. The standard normal, percentile, BC, and BCA, and 
recentered percentile CIs require similar sample sizes to recover $\mu$ at the 
nominal level for all combinations of $n$ and $\phi$. All methods seem to 
require a smaller sample to recover $\mu$ at the nominal rate when dealing with 
negative dependence versus positive dependence. For example, BC CIs recover 
$\mu$ for $n \geq 100$ when $\phi = -0.2$, but they only recover $\mu$ for 
$n \geq 800$ when $\phi = 0.2$. In addition, as a negative dependence gets 
stronger, holding everything else equal, coverage increases, which lead to the 
Student $t$ CI's aforementioned over-coverage. As a positive dependence gets 
stronger, holding everything else equal, coverage decreases, and a larger sample 
is necessary to recover~$\mu$. A possible explanation for this is that if a 
stationary series has a positive 
autocorrelation, the effective sample
size is decreased, whereas is a series has a negative autocorrelation, the
effective sample size is increased. Additionally, this seems to have a 
greater effect on the the estimation of the location parameter versus that
of the scale parameter or temporal dependence parameter.

@@

\subsection*{Exponential Marginal Distribution}
For an exponential marginal distribution, the empirical coverage rates for 
$\mu$, $\sigma_w$, %\eds{define $\sigma_w$} 
%\mc{added definitions for $\mu$ and $\sigma_w$ in the above paragraph} 
and 
the lag-1 autocorrelation coefficient, %\mc{replaced $\phi$ here, as I understand
%it, we are not using $\phi$ in reference to the temporal dependence of the 
%exponentially distributed series?}, 
as well as 95\%
confidence intervals of the real coverage are displayed in 
\textbf{Figures~\ref{fig:exp_mu1}--\ref{fig:exp_phi1}}. 
Additionally, a set of 50 CIs are displayed for each of the percentile, 
BC, and recentered percentile approaches for exponentially distributed
samples of $n \in \{800, 1600\}$ with lag-1 autocorrelation coefficient 0.355
($\phi = 0.4$) in Figure~\ref{fig:eri}.

@@

@@

One of the goals of this study was to provide some practical recommendations for 
necessary sample sizes when using block bootstrap to estimate parameters of 
serially dependent data. When using Student's $t$ intervals and $\phi$ is 
unknown, the results of this study suggest that an $n \geq 800$ may be necessary 
for common practice to estimate simpler parameters like $\mu$ or $\sigma_x$. If 
$\phi$ is already known, a lesser $n$ may be adequate to estimate these 
parameters. For lower absolute values of $\phi$, a smaller $n$ is sufficient 
when estimating $\sigma_x$. For negative values of $\phi$, coverage of all 
parameters is higher in general. In fact, Student's $t$ seems to over-cover 
$\mu$ for $\phi = - 0.4$ when using smaller sample sizes, meaning that if $\phi$ 
is known to have a very negative value, another CI may be required. However, in 
real world applications, $\phi$ is more commonly found to be positive, so 
Student's $t$ should still be used if $\phi$ is unknown. Lastly, to estimate 
$\phi$, $n \geq 100$ using the Student's $t$ method may be sufficient. Further
investigation may be necessary to see if there are other percentile-based 
interval corrections
that fix the coverage deterioration problem for $\phi$.

Although
we have only used serial dependences as strong as 0.4, we have established the 
trends as $|\phi|$ gets larger. When estimating $\mu$, we expect coverage rates 
to decrease as
$\phi$ approaches 1 - as $\phi$ approaches -1, we may observe increased 
over-coverage. When estimating $\sigma$, we expect a larger sample size to
be necessary as $|\phi|$ gets closer to 1. Lastly, when estimating $\phi$ for
a marginal normal distribution, we 
expect a larger sample size to be necessary as $|\phi|$ approaches 0. However, 
for a marginal exponential distribution, we expect coverage rates of the 
first-order autocorrelation to 
respond
to stronger dependences in a trend similar to that of coverage rates of $\mu$.

@@

\textbf{Figures~\ref{fig:mu2}--\ref{fig:phi2}} 
summarizes the empirical coverage rates and the 95\% confidence intervals of the 
real coverage for a standard normal marginal distribution using block bootstrap
with $l = \lceil 2n^{1/3} \rceil$.

@@

\textbf{Figure~\ref{fig:sigma2}} looks very similar to 
\textbf{Figure~\ref{fig:sigma1}}, but coverage rates of $\sigma$ do look 
slightly lower
especially for negative values of $\phi$, although Student's $t$ CIs are again
not as influenced by this change in $l$. A larger sample size seems necessary
when using other CIs to estimate $\sigma$ for $l = \lceil 2n^{1/3} \rceil$.

@@

\textbf{Figures~\ref{fig:exp_mu2}--\ref{fig:exp_phi2}} 
summarizes the empirical coverage rates and the 95\% confidence intervals of the 
real coverage for a exponential marginal distribution using block bootstrap
with $l = \lceil 2n^{1/3} \rceil$.

@@

For a stationary series with a marginal exponential distribution, altering the 
block size results in the same 
changes in performance of different CIs.

>>>>

\subsection*{Marginal Standard Normal Distribution}
%We designed a simulation study to compare the performance of the different block 
%bootstrap CI methods. In particular, 
We generated time series $X_t$ from a 1st 
order autoregressive (AR(1)) process:
\begin{equation*}
X_t = \phi X_{t-1} + \epsilon_t,
\end{equation*}
where $\phi$ is an autoregressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: $\{-0.4, -0.2, 0.0, 0.2, 0.4\}$. We only
used serial dependences as strong as 0.4, because we only seek to 
establish the 
general trend as the strength of the autocorrelation 
increases, and how it varies depending on the sign of the autocorrelation and 
the parameter of interest. The 
series $X_t$ has mean zero and variance 
$\sigma_x^2 = (1 - \phi^2) \sigma_{\epsilon}^2$, so for each value of $\phi$, we 
set $\sigma_{\epsilon}^2 = 1 / (1 - \phi^2)$ such that $\sigma_x^2 = 1$.

@@

The coverage rates of the CIs were used to compare the performance of CIs. Let
$\hat\theta_{L, r}$ and $\hat\theta_{U, r}$ be the lower and upper bound,
respectively, for the confidence interval constructed for each replicate
$r \in \{1, \ldots, R\}$, where $R$ is the number of replicates. Then
$cov = \#\{\hat\theta_{L, r} < \theta < \hat\theta_{U, r} \}/R$, for
$r \in \{1, \ldots, R\}$. If a CI method is valid, then the coverage rate is
expected to match the nominal level. Because it is unlikely for the coverage to 
exactly match the nominal level, we can construct a 95\% Wald-type CI of the 
coverage rate 
(which is an estimate of a proportion with $R = 10000$). If the proportion 0.95 
is included in the interval, the block bootstrap method is likely performing 
well. If all values in the interval are below 0.95, the results would suggest 
that the method either is providing inaccurate estimation, is underestimating 
the process' variability, or a combination of both. If all values in the 
interval are above .95, the results suggest that the method is overestimating 
the process' variability. \citet{brown2001interval} notes that the Wald-type 
interval has poor coverage as the proportion approaches 0 or 1.
\textbf{Figures~\ref{fig:mu1}--\ref{fig:phi1}} 
summarizes the empirical coverage rates and the 95\% confidence intervals of the 
real coverage for a marginal standard normal distribution using block bootstrap
with $l = \lceil n^{1/3} \rceil$, generated using the 
R \textsl{ggplot2} package.\citep{ggplot2}

\subsection*{Marginal Unit Exponential Distribution}
Additionally, we evaluated the performance of block bootstrap when used to 
estimate
the mean, standard deviation, and the lag-1 autocorrelation coefficient of a
stationary series with a marginal exponential distribution. The motivation of
for this study was to investigate if the results are robust to nonnormal
marginal distributions.

@@

For a marginal exponential  distribution, the empirical coverage rates for 
$\mu$, $\sigma_w$, %\eds{define $\sigma_w$} 
%\mc{added definitions for $\mu$ and $\sigma_w$ in the above paragraph} 
and 
the lag-1 autocorrelation coefficient using block bootstrap
with $l = \lceil n^{1/3} \rceil$, %\mc{replaced $\phi$ here, as I understand
%it, we are not using $\phi$ in reference to the temporal dependence of the 
%exponentially distributed series?}, 
as well as 95\%
confidence intervals of the real coverage are displayed in 
\textbf{Figures~\ref{fig:exp_mu1}--\ref{fig:exp_phi1}}.

@@

\subsection*{Marginal Standard Normal Distribution}

@@

For estimating the mean parameter~$\mu$, \textbf{Figure~\ref{fig:mu1}} suggests 
that all methods eventually approach correct coverage of $\mu$ as sample size 
increases. Student's $t$ CIs appear to need the smallest sample size to achieve 
correct coverage, except for samples with strong negative dependence, in which 
case, they actually over-cover $\mu$ for smaller sample sizes. For instance, for 
a sample with $n = 100$ and $\phi = -0.4$, the lower bound for a Students $t$ 
CI's coverage of $\mu$ is greater than 95\%, whereas the coverage intervals for 
other methods contain 95\%. The standard normal, percentile, BC, and BCA, and 
recentered percentile CIs require similar sample sizes to recover $\mu$ at the 
nominal level for all combinations of $n$ and $\phi$. All methods seem to 
require a smaller sample to recover $\mu$ at the nominal rate when dealing with 
negative dependence versus positive dependence. For example, BC CIs recover 
$\mu$ for $n \geq 100$ when $\phi = -0.2$, but they only recover $\mu$ for 
$n \geq 800$ when $\phi = 0.2$. In addition, as a negative dependence gets 
stronger, holding everything else equal, coverage increases, which lead to the 
Student $t$ CI's aforementioned over-coverage. As a positive dependence gets 
stronger, holding everything else equal, coverage decreases, and a larger sample 
is necessary to recover~$\mu$. A possible explanation for this is that if a 
stationary series has a positive 
autocorrelation, the effective sample
size is decreased, whereas is a series has a negative autocorrelation, the
effective sample size is increased. Additionally, this seems to have a 
greater effect on the the estimation of the location parameter versus that
of the scale parameter or temporal dependence parameter.

@@

\subsection*{Marginal Unit Exponential Distribution}

@@

Additionally, a set of 50 CIs are displayed for each of the percentile, 
BC, and recentered percentile approaches for exponentially distributed
samples of $n \in \{800, 1600\}$ with lag-1 autocorrelation coefficient 0.355
($\phi = 0.4$) in Figure~\ref{fig:eri}.

@@

One of the goals of this study was to provide some practical recommendations for 
necessary sample sizes when using block bootstrap to estimate parameters of 
serially dependent data. When using Student's $t$ intervals and the marginal
distribution and temporal dependence is 
unknown, the results of this study suggest that $n \geq 1600$ may be necessary 
for common practice to estimate $\mu$, whereas $n > 3200$ may be necessary
to estimate the standard deviation. If 
the dependences tructure or the marginal distribution is already known, a 
lesser $n$ may be 
adequate to estimate these 
parameters. For lower absolute values of $\phi$, a smaller $n$ is sufficient 
when estimating the standard deviation. For negative values of $\phi$, coverage of all 
parameters is higher in general. Under a marginal standard normal distribution, 
Student's $t$ seems to over-cover 
$\mu$ for $\phi = - 0.4$  when 
using smaller sample sizes, meaning that if $\phi$ 
is known to have a very negative value, another CI may be required. However, in 
real world applications, $\phi$ is more commonly found to be positive, so 
Student's $t$ should still be used if $\phi$ is unknown. 
Lastly, to estimate 
$\phi$, $n \geq 100$ using the Student's $t$ method may be sufficient under
a marginal standard normal distribution, whereas $n \geq 1600$ may be required
under under a marginal exponential distribution. Further
investigation may be necessary to see if there are other percentile-based 
interval corrections
that fix the coverage deterioration problem for $\phi$.

Although
we have only used serial dependences as strong as 0.4, we have established the 
trends as $|\phi|$ gets larger. When estimating $\mu$, we expect coverage rates 
to decrease as
$\phi$ approaches 1 - as $\phi$ approaches -1, we may observe increased 
over-coverage. When estimating the standard deviation, we expect a larger sample 
size to
be necessary as $|\phi|$ gets closer to 1. Lastly, when estimating $\phi$ for
a marginal normal distribution, we 
expect a larger sample size to be necessary as $|\phi|$ approaches 0. However, 
for a marginal exponential distribution, we expect coverage rates of the 
first-order autocorrelation to 
respond
to stronger dependences in a trend similar to that of coverage rates of $\mu$.

@@

\textbf{Figures~\ref{fig:mu2}--\ref{fig:phi2}} 
summarizes the empirical coverage rates and the 95\% confidence intervals of the 
real coverage for a marginal standard normal distribution using block bootstrap
with $l = \lceil 2n^{1/3} \rceil$.

@@

\textbf{Figure~\ref{fig:sigma2}} looks very similar to 
\textbf{Figure~\ref{fig:sigma1}}, but coverage rates of $\sigma_x$ do look 
slightly lower
especially for negative values of $\phi$, although Student's $t$ CIs are again
not as influenced by this change in $l$. A larger sample size seems necessary
when using other CIs to estimate $\sigma_x$ for $l = \lceil 2n^{1/3} \rceil$.

@@

\textbf{Figures~\ref{fig:exp_mu2}--\ref{fig:exp_phi2}} 
summarizes the empirical coverage rates and the 95\% confidence intervals of the 
real coverage for a marginal exponential distribution using block bootstrap
with $l = \lceil 2n^{1/3} \rceil$.

@@

For a stationary series with a marginal unit exponential distribution, altering 
the 
block size results in the same 
changes in performance of different CIs.
